---
title: "Break Free From Plastic"
description: |
  In early January, I participated in #TidyTuesday. Here's the creative plot I made and how I made it.
author:
  - name: Mukta Kelkar
    url: {}
date: 03-13-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(jpeg)
library(grid)
library(ggpubr)
library(packcircles)

```

## Step 1: Read in and wrangle data

First, I read in and wrangled data from the TidyTuesday [master github page](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-01-26#get-the-data-here). The data are from the organization [Break Free From Plastic](https://www.breakfreefromplastic.org/). 

```{r}
#read in data
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv') %>% 
  #make data tidy
  pivot_longer(empty:grand_total, 
               names_to = 'plastic_type', 
               values_to = 'count',
               values_drop_na = TRUE) %>% 
  filter(count != 0)

plastics_ps <- plastics %>% 
  select(-c(num_events, volunteers, year)) %>%   #remove num_events and volunteers columns
  filter(!(parent_company %in% c("Grand Total", "null", "Null", "Unbranded", "NULL", "Assorted"))) %>% #remove rows with unknown company names
  filter(plastic_type == "ps") %>%   #only include polystyrene
  filter(!(country == "EMPTY")) %>%   #drop rows with no country name
  group_by(country) %>% 
  summarize(
    total = sum(count)
  )   #create a summary table of amount of polystyrene found in each country

```

## Step 2: Make a bubble plot

In the spirit of #TidyTuesday, I set out to make a fun graph I'd never made before, so I made a bubble plot using the `packcircles` package. I then followed the tutorial posted in [R Graph Gallery](https://www.r-graph-gallery.com/306-custom-circle-packing-with-one-level.html)

The image I used as my background is from [The Huffington Post](https://www.huffingtonpost.co.uk/entry/living-near-landfill-site-increases-cancer-risk_uk_574569d4e4b03e9b9ed4fa19?guccounter=1&guce_referrer=aHR0cHM6Ly9kdWNrZHVja2dvLmNvbS8&guce_referrer_sig=AQAAABC-T6g5EMhu6-U0NXBHYAGyQDYx9YCejy7aNJi_WzIekAHvwxEf2hhi4dA7sWuOrDEZ9xu5_rZare35FJ8ECgCf5LdYKwALVDh0h2c3QADvkw4lx9MqOHVgP0fIDyfqWl_DeUYmN3i9d36QR-piFpqkxJzOj78Em3D5LfYsEZjc).

```{r}
#see what sample data from instructions looks like
data <- data.frame(group=paste("Group", letters[1:20]), value=sample(seq(1,100),20))  #wooo its the same format as my pastics data thank god

# Generate the layout. This function return a dataframe with one line per bubble. 
# It gives its center (x and y) and its radius, proportional of the value
packing <- circleProgressiveLayout(plastics_ps$total, sizetype = 'area')

# We can add these packing information to the initial data frame
plastic_circle <- cbind(plastics_ps, packing)

# Check that radius is proportional to value. We don't want a linear relationship, since it is the AREA that must be proportionnal to the value
#plot(plastic_circle$radius, plastic_circle$total)

# The next step is to go from one center + a radius to the coordinates of a circle that
# is drawn by a multitude of straight lines.
dat.gg <- circleLayoutVertices(packing, npoints=50)

dat.gg$total <- rep(plastics_ps$total, each = 51)

#read in image
landfill <- jpeg::readJPEG("landfill.jpg")


# Make the plot
ggplot() + 
  #Make the bubles
  background_image(landfill) +
  geom_polygon(data = dat.gg, 
               aes(x, y, group = id, fill = as.factor(id)), alpha = 0.95) +
  
  #Add text in the center of each bubble + control its size
  geom_text(data = plastic_circle, aes(x, y, size = total, label = country)) +
  scale_size_continuous(range = c(1,4)) +
  
  #General theme
  #scale_colour_viridis_b() +
  theme_void() +
  theme(legend.position = "none") +
  coord_equal() +
  labs(title = "Polystyrene (aka Stryrofoam) found in countries around the world",
       subtitle = "Data collected by #BreakFreeFromPlastic")


```



---
Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


