---
title: "Oil Spill Spatial Analysis"
description: |
  Another example of my coding skills I'm excited to share are my spatial maping skills using shapefiles. The following is also an assignment from my advanced data analysis course.
author:
  - name: Mukta Kelkar
    url: {}
date: 02-24-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(sf)
library(janitor)
library(tmap)
library(sp)
```

### Step 1: Read in data

First, I use the `sf` and `here` packages to read in shapefiles of the State of California and Oil Spill Incident Tracking data. I then use `st_transform` to make sure both shapefiles have the same CRS. 

```{r read in data, cache = TRUE, message = FALSE, warning = FALSE, results='hide'}
#Read in oil spill layer
ds394 <- st_read("ds394", layer = "ds394") %>% 
  clean_names()

#Check the projection:
#st_crs(ds394) #NAD83

# Read in the CA county data (TIGER shapefile):
ca_counties <- st_read("ca_counties", layer = "CA_Counties_TIGER2016") %>% 
  clean_names()

#Check the projection:
#st_crs(ca_counties) #WGS 84

#Make ca_counties match the projection of ds394
ca_counties <- st_transform(ca_counties, st_crs(ds394))

#confirm that the projection is updated
#st_crs(ca_counties) #NAD83 confirmed

#Make a quick exploratory ggplot
#ggplot() +
  #geom_sf(data = ca_counties) +
  #geom_sf(data = ds394)
```

### Step 2: Make an Interactive Map

I used `tmap` to create an exploratory interactive map of oil spills in California.

```{r interactive map, message = FALSE, warning = FALSE}
#Set tmap mode to interactive viewing
tmap_mode("view")

#Create exploratory interactive map
tm_shape(ds394) +
  tm_dots("localecoun") #dots are colored based on the county the spill occured in
```

### Step 3: Make a chloropleth map

Here I use `ggplot2` and `sf` to create a chloropleth map to visualize the number of inland oil spill events in every county in California.

```{r cloropleth map, message = FALSE, warning = FALSE}
#assume no duplicate oil spill events

#make a subset of relevant ca county data
ca_subset <- ca_counties %>% 
  select(name) %>%     #select variable of interest
  rename(county_name = name) #rename variable

#make a subset of relevant oil spill data
oil_subset <- ds394 %>% 
  select(localecoun, inlandmari) %>% #select variables of interest
  filter(inlandmari == "Inland") %>%  #filter dataset to only include inland observations
  rename(county_name = localecoun)  #rename county column to match ca_subset

#Join the oil spill and ca county datasets by variable `county_name`
ca_oil <- ca_subset %>% 
  st_join(oil_subset)

#Find count of oil spills observed by county. 
oil_counts <- ca_oil %>% 
  count(county_name.y)

#Create chloropleth map using ggplot
ggplot(data = oil_counts) +
  geom_sf(aes(fill = n), color = "gray100", size = 0.05) +
  scale_fill_gradientn(colors = c("gray90", "gray0")) +
  theme_minimal() +
  labs(fill = "Number of Oil Spills",
       title = "Inland Oil Spills in California Counties",
       caption = "Los Angeles County has the most oil spills out of all the CA Counties.")
  
```

### References

California Department of Fish and Game, Office of Spill Prevention and Response. (2009). Oil Spill Incident Tracking [ds394]. California Department of Fish and Game. Available at: [https://map.dfg.ca.gov/metadata/ds0394.html](https://map.dfg.ca.gov/metadata/ds0394.html)   

US Census Bureau, Department of Commerce. (2016). CA Counties TIGER Line Shapefile.


---

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


