---
title: "Working with Rasters"
description: |
  Last post I wanted to share my coding skills with shapefiles, and today I want to share my skills working wtih rasters. In the following post, I wrangle cetacean probability raster data from AquaMaps to create a map of cetacean species richness off the coast of California.
author:
  - name: Mukta Kelkar
    url: {}
date: 02-25-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(raster)  
library(tidyverse)
library(here)
library(rnaturalearth)
```

### Step 1: Read in Data

First I read in all the CA cetacean .tif files from AquaMaps and save them as a RasterStack using `raster`. The data shows the probability of occurrence of 35 cetacean species that can be found along the California coast.  I also read in a shapefile of countries of the world using `rnaturalearth`.

```{r}
#Pull all .tif files from folder
my_files <- list.files(path = "ca_cetaceans", full.names = TRUE)

#Use `raster::stack` to rasterize all the layers at once
cetacean_stack <- raster::stack(my_files)

#Read in shapefile using `rnaturalearth`
coastline <- ne_countries(scale = 110, returnclass = "sf")

#View data
#cetacean_stack
#summary(cetacean_stack)

#Preliminary visualization of data
#plot(cetacean_stack)

```

### Step 2: Filer the Data

Next I filter the probability data using a manually created filter function. I filter the data so that if the probability of occurrence is greater than or equal to `0.75`, the cetacean species is logged as being present in that raster square. I then add up all the filtered rasters to create one RasterLayer.

```{r}
#Create a filter function 
is_present <- function(x, thresh = 0.75) {
    y <- ifelse(x >= thresh, 1, NA)
    return(y)
}

#filter the raster
cetacean_present <- calc(cetacean_stack, fun = is_present)

#add up all the raster layers
cetacean_prob <- calc(cetacean_present, fun = sum, na.rm = TRUE)

#change the initial raster into a dataframe
cetacean_stack_df <- raster::rasterToPoints(cetacean_stack) %>%
  as.data.frame()

#change the filtered raster layer into a dataframe
cetacean_prob_df <- raster::rasterToPoints(cetacean_prob) %>%
  as.data.frame()

#cetecean_prob_df[is.na(cetecean_prob_df[])] <- 0
```

### Step 3: Create a Static Map

Lastly I use `ggplot2` and `rnaturalearth` to create a map of the cetacean species richness along the coast of California.

```{r}
ggplot() +
  geom_raster(data = cetacean_prob_df, aes(x = x, y = y, fill = layer)) +
  geom_sf(data = coastline) +
  coord_sf(expand = 0, xlim = c(-130,-110), ylim = c(28,45)) +
  scale_fill_gradient2(low = 'yellow', mid = "red", high =  'purple', midpoint = 13) +
  labs(
    x = "Latitude",
    y = "Longitude",
    fill = "Species Richness",
    title = "Cetacean species richness along the California coast",
    caption = "Fig. 1: Cetacean species richness calculated using a 0.75 probability threshold."
  ) +
  theme_minimal()
  
```


---
Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


